Towards Patient-Specific Neural Field Modeling using Micro-Electrode Data

### 2011-04-26 ltx/data_paper/DataPaper.tex
added bib file to data paper folder
added xi estiamation to data paper

\ifpdf
\usepackage[pdftex]{graphicx}
\else
\usepackage{graphicx}
\fi
\usepackage{color}
\newcommand{\dean}[1]{\textsf{\emph{\textbf{\textcolor{red}{#1}}}}} 
\newcommand{\red}{\textcolor{red}}
\newcommand{\blue}{\textcolor{blue}}
\newcommand{\cyan}{\textcolor{cyan}}

@@

Substituting  \ref{eq:Qfunctiontrace1} and \ref{eq:Qfunctiontrace2} into \ref{eq:QIntermsofJointDist} and taking the expectation of the log likelihood function for all time instants gives the lower-bound that needs to be maximised to gain optimal parameter estimates,
\begin{align}\label{eq:Voldermont}
 \mathcal Q(\boldsymbol \Theta,\boldsymbol\Theta')&= -(T-1)n_y\ln \sigma_{\epsilon}^2-\frac{1}{\sigma_{\epsilon}^2}\mathrm{tr}\left\lbrace\boldsymbol\sum_{t=0}^{T-1}\left[ (\mathbf y_{t+1}-\mathbf C\mathbf{\hat{x}}_{t+1}) (\mathbf y_{t+1}-\mathbf C\mathbf{\hat{x}}_{t+1})^\top+\mathbf C \mathbf P_{t+1}\mathbf C^\top\right] \right\rbrace\nonumber \\
&-(T-1)n_x\ln\sigma_d^2-\frac{1}{\sigma_d^2}\mathrm{tr}\left\lbrace \mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_{t+1}\mathbf x_{t+1}^\top\right]\tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace +\frac{2\xi}{\sigma_d^2} \mathrm{tr}\left\lbrace \mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_t\mathbf x_{t+1}^\top\right] \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace \nonumber \\
&-\frac{\xi^2}{\sigma_d^2}\mathrm{tr} \left\lbrace\mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_t\mathbf x_{t}^\top\right]\tilde{\boldsymbol\Sigma}_e^{-1} \right\rbrace +\frac{2}{\sigma_d^2}\mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_{t+1}^\top\tilde{\boldsymbol\Sigma}_e^{-1}q( \mathbf x_t)\right]\boldsymbol\theta
 \nonumber \\
&-\frac{1}{\sigma_d^2}\boldsymbol\theta^\top \mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}  q^\top(\mathbf  x_t)\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf x_t)\right]\boldsymbol\theta-\frac{2\xi}{\sigma_d^2} \mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1} \mathbf x_t^\top\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf x_t)  \right] \boldsymbol\theta
\end{align}
It should be noted that the second term in equation \ref{eq:Voldermont} is obtained by adding and subtracting $\mathbf C \mathbf{\hat{x}}_{t+1} \mathbf{\hat{x}}_{t+1}^\top \mathbf C^\top$ to the expectation, $\mathbf E_{\boldsymbol\Theta'}\left[ .\right]$, ref(Shumway). By choosing
\begin{align}
	\Xi_{0}&=\mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_{t+1}\mathbf x_{t+1}^\top\right]\label{eq:defofXi0} \\
\Xi_{1}&=\mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_t\mathbf x_{t+1}^\top\right]\label{eq:defofXi1} \\
\Xi_{2}&=\mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_t\mathbf x_{t}^\top\right]\\
\Xi_{3}&=\mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_{t+1}^\top\tilde{\boldsymbol\Sigma}_e^{-1}q( \mathbf x_t)\right] \label{eq:Xi3} \\	
\Xi_{4}&= \mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}  q^\top(\mathbf  x_t)\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf x_t)\right] \label{eq:Xi4}\\
 \Xi_{5}&=\mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1} \mathbf x_t^\top\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf x_t)  \right] \label{eq:Xi5}
 \end{align}
equation \ref{eq:Voldermont} can be written as
\begin{align}\label{eq:VoldermontwitXivariables}
 \mathcal Q(\boldsymbol \Theta,\boldsymbol\Theta')&=-(T-1)n_y\ln \sigma_{\epsilon}^2-\frac{1}{\sigma_{\epsilon}^2}\mathrm{tr}\left\lbrace\boldsymbol\sum_{t=0}^{T-1}\left[ (\mathbf y_{t+1}-\mathbf C\mathbf{\hat{x}}_{t+1}) (\mathbf y_{t+1}-\mathbf C\mathbf{\hat{x}}_{t+1})^\top+\mathbf C \mathbf P_{t+1}\mathbf C^\top\right] \right\rbrace\nonumber \\
&-(T-1)n_x\ln\sigma_d^2-\frac{1}{\sigma_d^2}\mathrm{tr}\left\lbrace \Xi_{0} \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace +\frac{2\xi}{\sigma_d^2} \mathrm{tr}\left\lbrace \Xi_{1} \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace -  \frac{\xi^2}{\sigma_d^2}\mathrm{tr} \left\lbrace\Xi_{2}\tilde{\boldsymbol\Sigma}_e^{-1} \right\rbrace 
 \nonumber \\
&+\frac{2}{\sigma_d^2}\Xi_{3}\boldsymbol\theta-\frac{1}{\sigma_d^2}\boldsymbol\theta^\top \Xi_{4}\boldsymbol\theta-\frac{2\xi}{\sigma_d^2} \Xi_{5} \boldsymbol\theta.
\end{align}

Consider first the maximization of $\mathcal{Q}$-function with respect to $\boldsymbol\theta$, we can write
\begin{equation}\label{eq:Qpartialtheta}
 \frac{\partial \mathcal Q}{\partial \boldsymbol\theta}=\frac{2}{\sigma_d^2}\Xi_3-\frac{1}{\sigma_d^2}\boldsymbol\theta^\top(\Xi_4^\top+\Xi_4)-\frac{2\xi}{\sigma_d^2}\Xi_5,
\end{equation}
variable $\Xi_4$ is symmetric and positive-definite by construction, equating \ref{eq:Qpartialtheta} to the zero vector yields
\begin{equation}\label{eq:theta}
 \boldsymbol \theta=\Xi_4^{-1}\left(\Xi_3^\top-\xi\Xi_5^\top \right).
\end{equation}
The second derivative is negative definite at this stationary point, and hence it is a global maximizer of $\mathcal{Q}$-function.
Now maximizing with respect to $\sigma_{\epsilon}^2$ we get
\begin{equation}
  \frac{\partial \mathcal Q}{\partial \sigma_{\epsilon}^2}=-(T-1)n_y\frac{1}{\sigma_{\epsilon}^2}+\frac{1}{\sigma_{\epsilon}^4}\mathrm{tr}\left\lbrace\sum_{t=0}^{T-1}\left[ (\mathbf y_{t+1}-\mathbf C\mathbf{\hat{x}}_{t+1}) (\mathbf y_{t+1}-\mathbf C\mathbf{\hat{x}}_{t+1})^\top+\mathbf C \mathbf P_{t+1}\mathbf C^\top\right] \right\rbrace
\end{equation}
setting above to zero and solving for $\sigma_{\epsilon}^2$ gives
\begin{equation}
 \sigma_{\epsilon}^2=\frac{1}{(T-1)n_y}\sum_{t=0}^{T-1}\mathrm{tr}\left\lbrace (\mathbf y_{t+1}-\mathbf C\mathbf{\hat{x}}_{t+1}) (\mathbf y_{t+1}-\mathbf C\mathbf{\hat{x}}_{t+1})^\top+\mathbf C \mathbf P_{t+1}\mathbf C^\top \right\rbrace
\end{equation}
A second derivative test shows that this is indeed the maximum. Finally differentiating with respec to $\sigma_d^2$ we get
\begin{align}
 \frac{\partial \mathcal Q}{\partial \sigma_d^2}&= -(T-1)n_x\frac{1}{\sigma_d^2}+\frac{1}{\sigma_d^4}\mathrm{tr}\left\lbrace \Xi_0 \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace-\frac{2\xi}{\sigma_d^4} \mathrm{tr}\left\lbrace \Xi_1 \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace+\frac{\xi^2}{\sigma_d^4}\mathrm{tr} \left\lbrace\Xi_2\tilde{\boldsymbol\Sigma}_e^{-1} \right\rbrace  \nonumber \\
&-\frac{2}{\sigma_d^4}\Xi_3\boldsymbol\theta +\frac{1}{\sigma_d^4}\boldsymbol\theta^\top \Xi_4\boldsymbol\theta+\frac{2\xi}{\sigma_d^4} \Xi_5 \boldsymbol\theta 
\end{align}
setting above to zero and solving for $\sigma_e^2$ we have
\begin{align}
\sigma_d^2=\frac{1}{(T-1)n_x}\left[ \mathrm{tr}\left\lbrace \Xi_0 \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace-
2\xi\mathrm{tr}\left\lbrace \Xi_1 \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace +\xi^2\mathrm{tr} \left\lbrace\Xi_2\tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace -2\Xi_3\boldsymbol\theta+\boldsymbol\theta^\top \Xi_4\boldsymbol\theta+2\xi\Xi_5 \boldsymbol\theta\right]  
\end{align}
The second derivative test confirms this is a maximum.

@@

% all figures here
% ~~~~~~~~~~~~~~~~
\begin{figure}[!ht]
\begin{center}
\includegraphics{./Figures/LFPs.eps}
\end{center}
\caption{{\bf Local field potentials from the micro-electrode array}. Data recorded from the first 8 channels of the array. The red dotted line indicates the seizure onset and the blue dashed line indicates the seizure end.}
\label{fig:TimeSeries}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Figures/FieldObservations.eps}
\end{center}
\caption{{\bf Snap-shots of the spatial aspects of the observed field}. The snap-shots are ordered from A to I forming a consecutive sequence spaced 2.5~ms apart starting at 34.024s}
\label{fig:FieldObserations}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Figures/TemporalFreq.eps}
\end{center}
\caption{{\bf Spectra from local field potential recordings}. A. Pre-seizure observations. B. Seizure observations. C. Post-seizure observations.}
\label{fig:TemporalFreqObservation}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Figures/SpatialFreq.eps}
\end{center}
\caption{{\bf Spatial frequency analysis of the observed neural field}. A. Pre-seizure observations. B. Seizure observations. C. Post-seizure observations.}
\label{fig:SpatialFreqObservation}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Figures/SpatialFreqCrossSection.eps}
\end{center}
\caption{{\bf Diagonal cross-section of the spatial frequency plots of the observed neural field}. A. Pre-seizure observations. B. Seizure observations. C. Post-seizure observations.}
\label{fig:DiagSpatialFreqObservation}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Figures/CrossCorr2D.eps}
\end{center}
\caption{{\bf Mean cross-correlations between consecutive observations of the neural field}. Note: The correlations are normalised such that the maximum value is one. A. Pre-seizure observation correlations. B. Seizure observation correlations. C. Post-seizure observation correlations.}
\label{fig:SpatialCrossCorrelation}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Figures/HomoTestCrossCorr.eps}
\end{center}
\caption{{\bf Mean cross-correlations between consecutive samples using subsets of observations of the neural field}. Each subplot shows correlations corresponding to a subset electrodes. Indexing the electrodes in a (x,y) grid, where index ranges from 1 to 10. Subplot A shows the mean correlation between electrode indexes (1-6, 5-10), subplot B shows correlations from indexes (5-10,5-10), subplot C shows indexes (1-6, 1-6) and subplot D shows indexes (5-10, 5-10).}
\label{fig:HomogeneityTest}
\end{figure}

>>>>

% \ifpdf
% \usepackage[pdftex]{graphicx}
% \else
% \usepackage{graphicx}
% \fi
\usepackage{color}
\newcommand{\dean}[1]{\textsf{\emph{\textbf{\textcolor{red}{#1}}}}} 
\newcommand{\red}{\textcolor{red}}
\newcommand{\blue}{\textcolor{blue}}
\newcommand{\cyan}{\textcolor{cyan}}

@@

Substituting  \ref{eq:Qfunctiontrace1} and \ref{eq:Qfunctiontrace2} into \ref{eq:QIntermsofJointDist} and taking the expectation of the log likelihood function for all time instants gives the lower-bound that needs to be maximised to gain optimal parameter estimates,
\begin{align}\label{eq:Voldermont}
 \mathcal Q(\boldsymbol \Theta,\boldsymbol\Theta')&= -(T-1)n_y\ln \sigma_{\epsilon}^2-\frac{1}{\sigma_{\epsilon}^2}\mathrm{tr}\left\lbrace\boldsymbol\sum_{t=0}^{T-1}\left[ (\mathbf y_{t+1}-\mathbf C\mathbf{\hat{x}}_{t+1}) (\mathbf y_{t+1}-\mathbf C\mathbf{\hat{x}}_{t+1})^\top+\mathbf C \mathbf P_{t+1}\mathbf C^\top\right] \right\rbrace\nonumber \\
&-(T-1)n_x\ln\sigma_d^2-\frac{1}{\sigma_d^2}\mathrm{tr}\left\lbrace \mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_{t+1}\mathbf x_{t+1}^\top\right]\tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace +\frac{2\xi}{\sigma_d^2} \mathrm{tr}\left\lbrace \mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_t\mathbf x_{t+1}^\top\right] \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace \nonumber \\
&-\frac{\xi^2}{\sigma_d^2}\mathrm{tr} \left\lbrace\mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_t\mathbf x_{t}^\top\right]\tilde{\boldsymbol\Sigma}_e^{-1} \right\rbrace +\frac{2}{\sigma_d^2}\mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_{t+1}^\top\tilde{\boldsymbol\Sigma}_e^{-1}q( \mathbf x_t)\right]\boldsymbol\theta
 \nonumber \\
&-\frac{1}{\sigma_d^2}\boldsymbol\theta^\top \mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}  q^\top(\mathbf  x_t)\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf x_t)\right]\boldsymbol\theta-\frac{2\xi}{\sigma_d^2} \mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1} \mathbf x_t^\top\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf x_t)  \right] \boldsymbol\theta
\end{align}
It should be noted that the second term in equation \ref{eq:Voldermont} is obtained by adding and subtracting $\mathbf C \mathbf{\hat{x}}_{t+1} \mathbf{\hat{x}}_{t+1}^\top \mathbf C^\top$ to the expectation, $\mathbf E_{\boldsymbol\Theta'}\left[ .\right]$, ref(Shumway). By choosing
\begin{align}
	\boldsymbol\Xi_{0}&=\mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_{t+1}\mathbf x_{t+1}^\top\right]\label{eq:defofXi0} \\
\boldsymbol\Xi_{1}&=\mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_t\mathbf x_{t+1}^\top\right]\label{eq:defofXi1} \\
\boldsymbol\Xi_{2}&=\mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_t\mathbf x_{t}^\top\right]\\
\boldsymbol\Xi_{3}&=\mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_{t+1}^\top\tilde{\boldsymbol\Sigma}_e^{-1}q( \mathbf x_t)\right] \label{eq:Xi3} \\	
\boldsymbol\Xi_{4}&= \mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}  q^\top(\mathbf  x_t)\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf x_t)\right] \label{eq:Xi4}\\
 \boldsymbol\Xi_{5}&=\mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1} \mathbf x_t^\top\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf x_t)  \right] \label{eq:Xi5}
 \end{align}
equation \ref{eq:Voldermont} can be written as
\begin{align}\label{eq:VoldermontwitXivariables}
 \mathcal Q(\boldsymbol \Theta,\boldsymbol\Theta')&=-(T-1)n_y\ln \sigma_{\epsilon}^2-\frac{1}{\sigma_{\epsilon}^2}\mathrm{tr}\left\lbrace\boldsymbol\sum_{t=0}^{T-1}\left[ (\mathbf y_{t+1}-\mathbf C\mathbf{\hat{x}}_{t+1}) (\mathbf y_{t+1}-\mathbf C\mathbf{\hat{x}}_{t+1})^\top+\mathbf C \mathbf P_{t+1}\mathbf C^\top\right] \right\rbrace\nonumber \\
&-(T-1)n_x\ln\sigma_d^2-\frac{1}{\sigma_d^2}\mathrm{tr}\left\lbrace \boldsymbol\Xi_{0} \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace +\frac{2\xi}{\sigma_d^2} \mathrm{tr}\left\lbrace \boldsymbol\Xi_{1} \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace -  \frac{\xi^2}{\sigma_d^2}\mathrm{tr} \left\lbrace\boldsymbol\Xi_{2}\tilde{\boldsymbol\Sigma}_e^{-1} \right\rbrace 
 \nonumber \\
&+\frac{2}{\sigma_d^2}\boldsymbol\Xi_{3}\boldsymbol\theta-\frac{1}{\sigma_d^2}\boldsymbol\theta^\top \boldsymbol\Xi_{4}\boldsymbol\theta-\frac{2\xi}{\sigma_d^2}\boldsymbol \Xi_{5} \boldsymbol\theta.
\end{align}

Consider first the maximization of $\mathcal{Q}$-function with respect to $\boldsymbol\theta$, we can write
\begin{equation}\label{eq:Qpartialtheta}
 \frac{\partial \mathcal Q}{\partial \boldsymbol\theta}=\frac{2}{\sigma_d^2}\Xi_3-\frac{1}{\sigma_d^2}\boldsymbol\theta^\top(\boldsymbol\Xi_4^\top+\boldsymbol\Xi_4)-\frac{2\xi}{\sigma_d^2}\boldsymbol\Xi_5,
\end{equation}
variable $\boldsymbol\Xi_4$ is symmetric and positive-definite by construction, equating \eqref{eq:Qpartialtheta} to the zero vector yields
\begin{equation}\label{eq:theta}
 \boldsymbol \theta=\boldsymbol\Xi_4^{-1}\left(\boldsymbol\Xi_3^\top-\xi\boldsymbol\Xi_5^\top \right).
\end{equation}
The second derivative is negative definite at this stationary point, and hence it is a global maximizer of $\mathcal{Q}$-function. In order to maximise $\mathcal{Q}$-function with respect to $\xi$ we can write
\begin{equation}\label{eq:Qpartialxi}
 \frac{\partial \mathcal Q}{\partial\xi}=\frac{2}{\sigma_d^2}\mathrm{tr}\left\lbrace \boldsymbol\Xi_1 \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace -\frac{2\xi}{\sigma_d^2}\mathrm{tr}\left\lbrace \boldsymbol\Xi_2 \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace-\frac{2}{\sigma_d^2}\boldsymbol\Xi_5,
\end{equation}
setting \eqref{eq:Qpartialxi} to zero and solving for $\xi$
\begin{equation}
 \xi=\frac{\mathrm{tr}\left\lbrace \boldsymbol\Xi_1 \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace-\boldsymbol\Xi_5\boldsymbol\Xi_4^{-1}\boldsymbol\Xi_3^\top}{\mathrm{tr}\left\lbrace \boldsymbol\Xi_2 \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace-\boldsymbol\Xi_5\boldsymbol\Xi_4^{-1}\boldsymbol\Xi_5^{\top}}
\end{equation}

Now maximizing with respect to $\sigma_{\epsilon}^2$ we get
\begin{equation}
  \frac{\partial \mathcal Q}{\partial \sigma_{\epsilon}^2}=-(T-1)n_y\frac{1}{\sigma_{\epsilon}^2}+\frac{1}{\sigma_{\epsilon}^4}\mathrm{tr}\left\lbrace\sum_{t=0}^{T-1}\left[ (\mathbf y_{t+1}-\mathbf C\mathbf{\hat{x}}_{t+1}) (\mathbf y_{t+1}-\mathbf C\mathbf{\hat{x}}_{t+1})^\top+\mathbf C \mathbf P_{t+1}\mathbf C^\top\right] \right\rbrace
\end{equation}
setting above to zero and solving for $\sigma_{\epsilon}^2$ gives
\begin{equation}
 \sigma_{\epsilon}^2=\frac{1}{(T-1)n_y}\sum_{t=0}^{T-1}\mathrm{tr}\left\lbrace (\mathbf y_{t+1}-\mathbf C\mathbf{\hat{x}}_{t+1}) (\mathbf y_{t+1}-\mathbf C\mathbf{\hat{x}}_{t+1})^\top+\mathbf C \mathbf P_{t+1}\mathbf C^\top \right\rbrace
\end{equation}
A second derivative test shows that this is indeed the maximum. Finally differentiating with respec to $\sigma_d^2$ we get
\begin{align}
 \frac{\partial \mathcal Q}{\partial \sigma_d^2}&= -(T-1)n_x\frac{1}{\sigma_d^2}+\frac{1}{\sigma_d^4}\mathrm{tr}\left\lbrace \Xi_0 \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace-\frac{2\xi}{\sigma_d^4} \mathrm{tr}\left\lbrace \Xi_1 \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace+\frac{\xi^2}{\sigma_d^4}\mathrm{tr} \left\lbrace\Xi_2\tilde{\boldsymbol\Sigma}_e^{-1} \right\rbrace  \nonumber \\
&-\frac{2}{\sigma_d^4}\Xi_3\boldsymbol\theta +\frac{1}{\sigma_d^4}\boldsymbol\theta^\top \Xi_4\boldsymbol\theta+\frac{2\xi}{\sigma_d^4} \Xi_5 \boldsymbol\theta 
\end{align}
setting above to zero and solving for $\sigma_e^2$ we have
\begin{align}
\sigma_d^2=\frac{1}{(T-1)n_x}\left[ \mathrm{tr}\left\lbrace \Xi_0 \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace-
2\xi\mathrm{tr}\left\lbrace \Xi_1 \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace +\xi^2\mathrm{tr} \left\lbrace\Xi_2\tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace -2\Xi_3\boldsymbol\theta+\boldsymbol\theta^\top \Xi_4\boldsymbol\theta+2\xi\Xi_5 \boldsymbol\theta\right]  
\end{align}
The second derivative test confirms this is a maximum.

@@

% % all figures here
% % ~~~~~~~~~~~~~~~~
% \begin{figure}[!ht]
% \begin{center}
% \includegraphics{./Figures/LFPs.eps}
% \end{center}
% \caption{{\bf Local field potentials from the micro-electrode array}. Data recorded from the first 8 channels of the array. The red dotted line indicates the seizure onset and the blue dashed line indicates the seizure end.}
% \label{fig:TimeSeries}
% \end{figure}
% 
% 
% \begin{figure}[!ht]
% \begin{center}
% \includegraphics{./Figures/FieldObservations.eps}
% \end{center}
% \caption{{\bf Snap-shots of the spatial aspects of the observed field}. The snap-shots are ordered from A to I forming a consecutive sequence spaced 2.5~ms apart starting at 34.024s}
% \label{fig:FieldObserations}
% \end{figure}
% 
% \begin{figure}[!ht]
% \begin{center}
% \includegraphics{./Figures/TemporalFreq.eps}
% \end{center}
% \caption{{\bf Spectra from local field potential recordings}. A. Pre-seizure observations. B. Seizure observations. C. Post-seizure observations.}
% \label{fig:TemporalFreqObservation}
% \end{figure}
% 
% \begin{figure}[!ht]
% \begin{center}
% \includegraphics{./Figures/SpatialFreq.eps}
% \end{center}
% \caption{{\bf Spatial frequency analysis of the observed neural field}. A. Pre-seizure observations. B. Seizure observations. C. Post-seizure observations.}
% \label{fig:SpatialFreqObservation}
% \end{figure}
% 
% \begin{figure}[!ht]
% \begin{center}
% \includegraphics{./Figures/SpatialFreqCrossSection.eps}
% \end{center}
% \caption{{\bf Diagonal cross-section of the spatial frequency plots of the observed neural field}. A. Pre-seizure observations. B. Seizure observations. C. Post-seizure observations.}
% \label{fig:DiagSpatialFreqObservation}
% \end{figure}
% 
% \begin{figure}[!ht]
% \begin{center}
% \includegraphics{./Figures/CrossCorr2D.eps}
% \end{center}
% \caption{{\bf Mean cross-correlations between consecutive observations of the neural field}. Note: The correlations are normalised such that the maximum value is one. A. Pre-seizure observation correlations. B. Seizure observation correlations. C. Post-seizure observation correlations.}
% \label{fig:SpatialCrossCorrelation}
% \end{figure}
% 
% \begin{figure}[!ht]
% \begin{center}
% \includegraphics{./Figures/HomoTestCrossCorr.eps}
% \end{center}
% \caption{{\bf Mean cross-correlations between consecutive samples using subsets of observations of the neural field}. Each subplot shows correlations corresponding to a subset electrodes. Indexing the electrodes in a (x,y) grid, where index ranges from 1 to 10. Subplot A shows the mean correlation between electrode indexes (1-6, 5-10), subplot B shows correlations from indexes (5-10,5-10), subplot C shows indexes (1-6, 1-6) and subplot D shows indexes (5-10, 5-10).}
% \label{fig:HomogeneityTest}
% \end{figure}

% tables here
% ~~~~~~~~~~~
\renewcommand{\arraystretch}{1.7}
\begin{table}[!ht]
\caption{
\bf{Algorithm for the Unscented RTS Smoother}}
\label{tab:URTSSAlgorithm}
\begin{tabular}{|c|}\hline
\multicolumn{1}{|p{16cm}|}{\textbf{1.} Forward initialisation} \\ 
$\hat{\mathbf x}_0, \mathbf P_0$ \\
\hline
\multicolumn{1}{|p{16cm}|}{\textbf{2.} Forward iteration: for $t \in \left\lbrace 0,\cdots, T\right\rbrace $, calculate the sigma points $\mathcal X_{i,t}^f$ using equations \ref{eq:sigmapoints1}-\ref{eq:sigmapoints3} and propagate through equation~\ref{eq:QmatrixForSigmapoints}} \\
$\mathcal X_{i,t+1}^{f-}=Q(\mathcal X_{i,t}^f) \quad i=0, \dots, 2L$\\
\multicolumn{1}{|p{16cm}|}{Calculate the predicted state and the predicted covariance matrix} \\
$\hat{\mathbf x}_{t+1}^{f-}=\sum_{i=0}^{2L} W_i^{(m)}\mathcal X_{i,t+1}^{f-} $ \\
$\mathbf P_{t +1}^{f-}=\sum_{i=0}^{2L} W_i^{(c)}(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})^\top+\boldsymbol \Sigma_e$ \\ 
\multicolumn{1}{|p{16cm}|}{Compute the filter gain, the filtered state and the filtered covariance matrix using the standard Kalman filter update equations} \\
$\mathcal K_{t+1}=\mathbf P_{t +1}^{f-}\mathbf C ^\top(\mathbf C \mathbf P_{t +1}^{f-}\mathbf C ^\top+\boldsymbol \Sigma_{\varepsilon})^{-1}$\\
$\hat{\mathbf x}_{t+1}^{f}=\hat{\mathbf x}_{t+1}^{f-}+\mathcal K_{t+1}(\mathbf y_{t+1}-\mathbf C\hat{\mathbf x}_{t +1}^{f-}) $\\
$\mathbf P_{t+1}^f=(\mathbf I - \mathcal K_{t+1}\mathbf C)\mathbf P_{t +1}^{f-}$\\ 
\hline
\multicolumn{1}{|p{16cm}|}{\textbf{3.} Backward initialisation}\\
$\mathbf P_T^b= \mathbf P_T^f, \quad \hat{\mathbf x}^b_T= \hat{\mathbf x}^f_T$ \\
\hline
\multicolumn{1}{|p{16cm}|}{\textbf{4.} Backward iteration: for $t \in \left\lbrace T-1, \cdots, 0 \right\rbrace $ calculate the sigma points $\mathcal X_{i,t}^b$ and propagate them through equation \ref{eq:QmatrixForSigmapoints}}\\
$\mathcal X_{i,t+1}^{b-}=Q(\mathcal X_{i,t}^b) \quad i=0, \dots, 2L$\\
\multicolumn{1}{|p{16cm}|}{Calculate the predicted state and the predicted covariance matrix}\\
$\hat{\mathbf x}_{t+1}^{b-}=\sum_{i=0}^{2L} W_i^{(m)}\mathcal X_{i,t+1}^{b-}$\\
$\mathbf P_{t +1}^{b-}=\sum_{i=0}^{2L} W_i^{(c)}(\mathcal X_{i,t+1}^{b-}-\hat{\mathbf x}_{t +1}^{b-})(\mathcal X_{i,t+1}^{b-}-\hat{\mathbf x}_{t +1}^{b-})^\top+\boldsymbol \Sigma_e $\\
$\mathbf M_{t +1}^{b-}=\sum_{i=0}^{2L} W_i^{(c)}(\mathcal X_{i,t}^{b-}-\hat{\mathbf x}_{t}^{f})(\mathcal X_{i,t+1}^{b-}-\hat{\mathbf x}_{t+1}^{b-})^\top$\\
\multicolumn{1}{|p{16cm}|}{Compute the smoother gain, the smoothed state and the smoothed covariance matrix}\\
$\mathbf S_t=\mathbf M_{t +1}^{b-}\left[ \mathbf P_{t +1}^{b-}\right] ^{-1} $\\
$\hat{\mathbf x}_t^b=\hat{\mathbf x}_t^f+\mathbf S_t\left[\hat{\mathbf x}_{t+1}^{b}-\hat{\mathbf x}_{t+1}^{b-}\right]$\\
$\mathbf P_{t}^{b}=\mathbf P_{t}^{f}+\mathbf S_t\left[\mathbf P_{t+1}^{b}-\mathbf P_{t+1}^{b-} \right]\mathbf S_t^\top $\\
\multicolumn{1}{|p{16cm}|}{Additional step to calculate the smoothed cross covariance matrix \cite{Sarkka2008}} \\
$\mathbf M_{t +1}^{b}=\mathbf S_t\mathbf P_{t+1}^{b}$\\
\hline
\end{tabular}
\begin{flushleft}This table shows the steps in the unscented Rauch-Tung-Striebel smoother algorithm. The steps are iterated 11 times for our state estimation procedure. The least squares algorithm is run after each iteration to update the parameter estimates.
\end{flushleft}
\label{tab:UKFAlgorithm}
\end{table}
 \renewcommand{\arraystretch}{1}
\newpage
\bibliographystyle{plain}
\bibliography{EMIDE}
\end{document}
